Manuscript_Tables_LongCovid

### 2023-10-09 characterisation/Manuscript_Tables_Figures.Rmd
revision

@@

get_baseline_main_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    select(X, COVID.19.infection, Tested.negative..earliest) %>%
    full_join(cprd_data%>%
                select("X",
                       "COVID.infection.CPRD" = "COVID.19.infection"  ,
                       "Tested.negative.CPRD" = "Tested.negative..earliest"
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                            ifelse(X == "Vaccination status = first dose vaccination (%)",                "   First dose vaccination", X)))) 
  Table1 <- kable(Table[c(1:34),],
                  col.names = c("", 
                                "COVID-19\ninfection", 
                                "First SARS-CoV-2 negative test ",
                                "COVID-19\ninfection", 
                                "First SARS-CoV-2 negative test "
                  )) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    add_header_above(c(" ", "SIDIAP" = 2, "CPRD" = 2)) %>%
    pack_rows("Comorbidities",26, 34) 
  Table1
}

@@

# symptoms table
get_symptoms_main_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    left_join(cprd_data%>%
                select("X",
                       "COVID.infection.CPRD" = "COVID.19.infection"  ,
                       "First.infection.CPRD" = "First.infection",
                       "Reinfections.CPRD" = "Reinfections",
                       "Tested.negative.all.CPRD" = "Tested.negative..all",
                       "Tested.negative.CPRD" = "Tested.negative..earliest"
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                     ifelse(X == "Vaccination status = first dose vaccination (%)",
                                   "   First dose vaccination", X)))) 
  Table1 <- kable(Table[c(1,35:74),],
                  col.names = c("",
                                "COVID-19\ninfection", 
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests",
                                "First SARS-CoV-2 negative test ",
                                 "COVID-19\ninfection", 
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests",
                                "First SARS-CoV-2 negative test "
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
       pack_rows("Symptoms",3,27 ) %>%
    add_header_above(c("", "SIDIAP" = 5, "CPRD" = 5)) 
  Table1
}

get_rr_infection_negative_plot <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
    scale_x_log10(breaks = c(0.5, 0.8, 1,2, 4, 6),
                  labels = c("0.5","0.8", "1","2", "4", "6"),
                  limits = c(0.5, 8)) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_infection_negative_plot_facet <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    facet_grid(cols = vars(wave)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
    scale_x_log10(breaks = c(0.5, 0.8, 1,2, 4, 6),
                  labels = c("0.5","0.8", "1","2", "4", "6"),
                  limits = c(0.5, 8)) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_first_reinfection_plot <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, -relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
     scale_x_log10(breaks = c(0.6, 0.8, 1, 2),
                   labels = c("0.6", "0.8", "1", "2"),
                  limits = c(0.5, 2)
                  ) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_first_reinfection_plot_facet <- function(data){
plot <- data  %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, -relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    facet_grid(cols = vars(wave)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
     scale_x_log10(breaks = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, 2),
                   labels = c("0.1", "0.2", "0.4","0.6", "0.8", "1", "2"),
                  limits = c(0.04, 2)
                  ) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

@@

### Figure 1: Incidence rates of COVID-19 and long-COVID-19
```{r, fig.dim = c(8, 8), echo = FALSE}
IR_data <- rbind(Incidence_rates_Covid_LongCov_CPRD, Incidence_rates_Covid_LongCov_SIDIAP)
Proportion_data <- rbind(Proportion_LongCovid_90days_CPRD, Proportion_LongCovid_90days_SIDIAP)

data_IR_fig90 <- IR_data %>%
  filter(cohort != "Long COVID-19 (28 days)" ) %>%
  filter(!(is.na(n_events))) %>%
  mutate(sex= as.character(sex)) %>%
  mutate(incidence_start_date = as.Date(incidence_start_date)) %>%
  select(
    incidence_start_date, n_persons, n_events, person_years, 
    ir_100000_pys = incidence_100000_pys,  
    ir_100000_pys_95CI_lower = incidence_100000_pys_95CI_lower, 
    ir_100000_pys_95CI_upper = incidence_100000_pys_95CI_upper, 
    age_group, sex, cohort, database
  ) %>%
  rbind(Proportion_data %>%
          mutate(person_years = person_days/365) %>%
          select(-person_days) %>%
          filter(cohort=="COVID-19 infection") %>%
          mutate(cohort = "Long COVID-19 among those with COVID-19, 90 days")
  ) %>%
  mutate(age_group = ifelse(age_group == "18 to 150", "Overall", 
                            ifelse(age_group == "80 to 150", "\u226580", age_group))) %>%
  mutate(age_group = factor(age_group,
                            levels = c("Overall", "18 to 34", "35 to 49", "50 to 64", "65 to 79", "\u226580"))) %>%
      mutate(cohort = ifelse(cohort == "Long COVID-19 among those with COVID-19, 90 days", 
                           "Long COVID symptoms among those with COVID-19",
                           ifelse(cohort == "Long COVID-19 (90 days)", "Long COVID symptoms after COVID-19 among the general population", cohort))) %>%
  mutate(cohort = factor(cohort, levels = c(
    "COVID-19 infections",
     "Long COVID symptoms after COVID-19 among the general population",
     "Long COVID symptoms among those with COVID-19"
  )))

@@

### Figure 2A: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days, overall
```{r, fig.dim = c(8, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste", "Altered smell or taste",
                          ifelse(symptom =="Cognitive dysfunction brain fog", "Cognitive dysfunction",
                                 ifelse(symptom == "Fatigue malaise", "Fatigue or malaise",
                                        ifelse(symptom == "Muscle spams pain", "Muscle spasms or pain", symptom
                                        )))))

@@

@@

### Figure 2B: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days, by waves

```{r, fig.dim = c(24, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_90days_CPRD) %>%
  filter(wave != "overall") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste", "Altered smell or taste",
                          ifelse(symptom =="Cognitive dysfunction brain fog", "Cognitive dysfunction",
                                 ifelse(symptom == "Fatigue malaise", "Fatigue or malaise",
                                        ifelse(symptom == "Muscle spams pain", "Muscle spasms or pain", symptom
                                        ))))) %>%
  mutate(wave = factor(wave, levels = c("wild", "alpha", "delta", "omicron")))

gg.rr.covid_negative_both_90 <- get_rr_infection_negative_plot_facet(data = RR_data)
gg.rr.covid_negative_both_90

@@

### Figure 3: RR of long-COVID symptoms: first infection vs re-infection - 90 days
```{r, fig.dim = c(8, 8), echo = FALSE}
RR_data <- rbind(RR_first_infection_reinfections_90days_SIDIAP,
                  RR_first_infection_reinfections_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste",
                          "Altered smell or taste",
                    ifelse(symptom =="Cognitive dysfunction brain fog",
                           "Cognitive dysfunction",
                    ifelse(symptom == "Fatigue malaise",
                           "Fatigue or malaise",
                    ifelse(symptom == "Muscle spams pain",
                           "Muscle spasms or pain", symptom
                                        )))))

 gg.rr.first_reinfection_90 <-  get_rr_first_reinfection_plot(data = RR_data)
gg.rr.first_reinfection_90
```

@@

### Incidence Rates
#### 1. Table: Incidence Rates of COVID-19 & long COVID symptoms
```{r}

@@

plot_IR <- function(age_to_filter, data, brk_vec){ 
  plot <- data %>%
  filter(age_group %in% age_to_filter) %>%
    mutate(cohort = ifelse(cohort == "COVID-19 infections", "COVID-19 infections",
                           ifelse(cohort == "Long COVID symptoms after COVID-19 among the general population", "Long COVID among\nthe general population", "Long COVID among\nthose infected"))) %>%
  ggplot() +
#  geom_smooth(aes(x =incidence_start_date, y =  ir_100000_pys), method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
geom_line(aes(x =incidence_start_date,
              y =  ir_100000_pys,
              color = age_group), size = 1.2)+
#    geom_ribbon(aes(x =incidence_start_date, 
    # y =  ir_100000_pys, color = age_group,
    # ymin = ir_100000_pys_95CI_lower, ymax = ir_100000_pys_95CI_upper
    # ), alpha = 0.1) +
#scale_y_continuous(limits = c(0,NA)) +
scale_x_date(breaks = brk_vec,
 #date_breaks = "2 month",
               date_labels = ("%b %y"),
               expand = c(0,0)) +
scale_colour_manual(values = c("#90e0ef","#00b4d8","#0077b6","#023E8A", "#01022F"
  #"#000000","#0075c4","#5b8c5a","#d8973c","#dc392e","#d8973c","#dc392e"
  # "#bee9e8","#62b6cb",","#cae9ff","#5fa8d3"
                                     )) +
    theme_bw()+
    theme(
      #panel.spacing.x = unit(4, "mm"),
      strip.text.y.left =  element_text(angle = 0),
      legend.position="top",
      plot.margin = unit(c(0, 0, 0, 0), "cm"))+
  ylab("IR per 100,000 p-y") +
 facet_grid(cohort ~sex  , 
            scales = "free") +    
    xlab("Date of COVID-19 infection") +
        guides(color = guide_legend(title = "Age group"))
   plot         
  }

@@

brk_vec <- seq.Date(as.Date("2020-10-01"),
                    as.Date("2020-12-01"),by="1 month")

@@

### Relative Risks - infections vs first negative test 
#### 1. Table: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days
```{r}
get_RR <- function(sidiap_data, cprd_data) {
  RR_infections <- rbind(sidiap_data, cprd_data
                      ) %>%
    filter(wave == "overall") %>%
    select(-"wave") %>%
  mutate(name_1 = "COVID-19 infection") %>%
  mutate(relative_risk = paste0(sprintf(relative_risk, fmt = '%.2f'),
                                " [",
                                sprintf(low_ci, fmt = '%.2f'),
                                ";",
                                sprintf(up_ci, fmt = '%.2f'),
                                "]"
         )) %>%
    select (-prop_1, -prop_2, -low_ci, -up_ci, 
                        -name_1, -name_2, -total_pop_1, -total_pop_2,) %>%
    relocate(database, .before = symptom)  %>%
    arrange(desc(database), desc(relative_risk))

@@

#### 2. Table: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 28 days
```{r}
table_RR_28 <- get_RR(sidiap_data = RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  cprd_data = RR_new_covid_tested_negative_earliest_28days_CPRD)
table_RR_28
```

#### 3. Figure: RR of long COVID symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_28days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  mutate(symptom = ifelse(symptom == "Altered smell taste", "Altered smell or taste",
                          ifelse(symptom =="Cognitive dysfunction brain fog", "Cognitive dysfunction",
                                 ifelse(symptom == "Fatigue malaise", "Fatigue or malaise",
                                        ifelse(symptom == "Muscle spams pain", "Muscle spasms or pain", symptom
                                        )))))

@@

#### 2. Table: RR of long-COVID symptoms, first infection vs re-infection - 28 days
```{r}
table_RR_re_28 <- get_RR_numbers_re_infec(sidiap_data = RR_first_infection_reinfections_28days_SIDIAP,
                                       cprd_data = RR_first_infection_reinfections_28days_CPRD)
 table_RR_re_28
```

#### 3. Figure: RR of long-COVID symptoms, first infection vs re-infection - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}

>>>>

correctSymptomNames <- function(x, column = "symptom") {
  pairs <- c(
    "Altered smell taste" = "Altered smell or taste",
    "Cognitive dysfunction brain fog" = "Cognitive dysfunction",
    "Fatigue malaise" = "Fatigue or malaise",
    "Muscle spams pain" = "Muscle spasms or pain",
    "Depression wo recurrence" = "Depression",
    "Dyspnea" = "Dyspnoea"
  )
  for (k in seq_along(pairs)) {
    x <- x |>
      mutate(!!column := gsub(
        pattern = names(pairs)[k], 
        replacement = unname(pairs[k]), 
        x = .data[[column]]
      ))
  }
  return(x)
}

@@

get_baseline_main_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    select(X, COVID.19.infection, Tested.negative..earliest) %>%
    full_join(cprd_data%>%
                select("X",
                       "COVID.infection.CPRD" = "COVID.19.infection"  ,
                       "Tested.negative.CPRD" = "Tested.negative..earliest"
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                            ifelse(X == "Vaccination status = first dose vaccination (%)",                "   First dose vaccination", X)))) %>%
    correctSymptomNames(column = "X")
  Table1 <- kable(Table[c(1:34),],
                  col.names = c("", 
                                "COVID-19\ninfection", 
                                "First SARS-CoV-2 negative test ",
                                "COVID-19\ninfection", 
                                "First SARS-CoV-2 negative test "
                  )) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
    add_header_above(c(" ", "SIDIAP" = 2, "CPRD" = 2)) %>%
    pack_rows("Comorbidities",26, 34) 
  Table1
}

@@

# symptoms table
get_symptoms_main_table <- function(sidiap_data, cprd_data) {
  Table <- sidiap_data %>%
    left_join(cprd_data%>%
                select("X",
                       "COVID.infection.CPRD" = "COVID.19.infection"  ,
                       "First.infection.CPRD" = "First.infection",
                       "Reinfections.CPRD" = "Reinfections",
                       "Tested.negative.all.CPRD" = "Tested.negative..all",
                       "Tested.negative.CPRD" = "Tested.negative..earliest"
                ) %>% 
          mutate(X = ifelse(X == "Trimester = jan-apr 2021 (%)",
                            "   Jan-apr 2021"  ,
                     ifelse(X == "Vaccination status = first dose vaccination (%)",
                                   "   First dose vaccination", X)))) %>%
    correctSymptomNames(column = "X")
  Table1 <- kable(Table[c(1,35:74),],
                  col.names = c("",
                                "COVID-19\ninfection", 
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests",
                                "First SARS-CoV-2 negative test ",
                                 "COVID-19\ninfection", 
                                "First infection",
                                "Re-infections",
                                "SARS-CoV-2 negative tests",
                                "First SARS-CoV-2 negative test "
                  ),
                  row.names = FALSE) %>%  
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  fixed_thead = T
    ) %>%
       pack_rows("Symptoms",3,27 ) %>%
    add_header_above(c("", "SIDIAP" = 5, "CPRD" = 5)) 
  Table1
}

get_rr_infection_negative_plot <- function(data){
plot <- data  %>%
  correctSymptomNames() %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
    scale_x_log10(breaks = c(0.5, 0.8, 1,2, 4, 6),
                  labels = c("0.5","0.8", "1","2", "4", "6"),
                  limits = c(0.5, 8)) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_infection_negative_plot_facet <- function(data){
plot <- data  %>%
  correctSymptomNames() %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    facet_grid(cols = vars(wave)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
    scale_x_log10(breaks = c(0.5, 0.8, 1,2, 4, 6),
                  labels = c("0.5","0.8", "1","2", "4", "6"),
                  limits = c(0.5, 8)) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_first_reinfection_plot <- function(data){
plot <- data  %>%
  correctSymptomNames() %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, -relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
     scale_x_log10(breaks = c(0.6, 0.8, 1, 2),
                   labels = c("0.6", "0.8", "1", "2"),
                  limits = c(0.5, 2)
                  ) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

get_rr_first_reinfection_plot_facet <- function(data){
plot <- data  %>%
  correctSymptomNames() %>%
    ggplot(aes(
      x = relative_risk,
      y = reorder(symptom, -relative_risk),
      xmin = low_ci,
      xmax = up_ci,
      shape = database,
      color = database)) +
    geom_point(size = 2, position = position_dodge(width = 1)) +
    geom_errorbar(width = 0, size = 1, position = position_dodge(width = 1)) +
    facet_grid(cols = vars(wave)) +
    scale_y_discrete(expand=c(0,0), position = "left") +
     scale_x_log10(breaks = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, 2),
                   labels = c("0.1", "0.2", "0.4","0.6", "0.8", "1", "2"),
                  limits = c(0.04, 2)
                  ) +  
    scale_colour_manual(values = c(col1, col2)) +
    geom_vline(
      xintercept = 1, colour = "#000000",
      linetype = 1
    ) +
    mytheme +
    xlab("\nRR (with 95% CI)") +
    ylab("")  +
    guides(color = guide_legend(title = "Database"),
           shape = guide_legend(title = "Database")) +
    geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5, 5.5, 6.5, 7.5, 8.5, 9.5,10.5,11.5,
                     12.5,13.5,14.5,15.5,16.5,17.5,18.5,19.5,20.5,21.5,22.5,23.5),
                linetype = "dashed",
               color = "#D2D6DB"
    )
plot
}

@@

### Figure 1: Incidence rates of COVID-19 and Post-acute COVID-19 symptoms
```{r, fig.dim = c(8, 8), echo = FALSE}
IR_data <- rbind(Incidence_rates_Covid_LongCov_CPRD, Incidence_rates_Covid_LongCov_SIDIAP)
Proportion_data <- rbind(Proportion_LongCovid_90days_CPRD, Proportion_LongCovid_90days_SIDIAP)

data_IR_fig90 <- IR_data %>%
  filter(cohort != "Long COVID-19 (28 days)" ) %>%
  filter(!(is.na(n_events))) %>%
  mutate(sex= as.character(sex)) %>%
  mutate(incidence_start_date = as.Date(incidence_start_date)) %>%
  select(
    incidence_start_date, n_persons, n_events, person_years, 
    ir_100000_pys = incidence_100000_pys,  
    ir_100000_pys_95CI_lower = incidence_100000_pys_95CI_lower, 
    ir_100000_pys_95CI_upper = incidence_100000_pys_95CI_upper, 
    age_group, sex, cohort, database
  ) %>%
  rbind(Proportion_data %>%
          mutate(person_years = person_days/365) %>%
          select(-person_days) %>%
          filter(cohort=="COVID-19 infection") %>%
          mutate(cohort = "Long COVID-19 among those with COVID-19, 90 days")
  ) %>%
  mutate(age_group = ifelse(age_group == "18 to 150", "Overall", 
                            ifelse(age_group == "80 to 150", "\u226580", age_group))) %>%
  mutate(age_group = factor(age_group,
                            levels = c("Overall", "18 to 34", "35 to 49", "50 to 64", "65 to 79", "\u226580"))) %>%
      mutate(cohort = ifelse(cohort == "Long COVID-19 among those with COVID-19, 90 days", 
                           "Post-acute COVID-19 symptoms among those with COVID-19",
                           ifelse(cohort == "Long COVID-19 (90 days)", "Post-acute COVID-19 symptoms after COVID-19 among the general population", cohort))) %>%
  mutate(cohort = factor(cohort, levels = c(
    "COVID-19 infections",
     "Post-acute COVID-19 symptoms after COVID-19 among the general population",
     "Post-acute COVID-19 symptoms among those with COVID-19"
  )))

@@

### Figure 2A: RR of Post-acute COVID-19 symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days, overall
```{r, fig.dim = c(8, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  correctSymptomNames()

@@

ggsave(file="Figure2.svg", plot=gg.rr.covid_negative_both_90, width=10, height=8)

@@

### Figure 2B: RR of Post-acute COVID-19 symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days, by waves

```{r, fig.dim = c(24, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_90days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_90days_CPRD) %>%
  filter(wave != "overall") %>%
  correctSymptomNames() %>%
  mutate(wave = factor(wave, levels = c("wild", "alpha", "delta", "omicron")))

gg.rr.covid_negative_both_90 <- get_rr_infection_negative_plot_facet(data = RR_data)
gg.rr.covid_negative_both_90 
ggsave(file="Figure3.svg", plot=gg.rr.covid_negative_both_90, width=10, height=8)

@@

### Figure 3: RR of Post-acute COVID-19 symptoms: first infection vs re-infection - 90 days
```{r, fig.dim = c(8, 8), echo = FALSE}
RR_data <- rbind(RR_first_infection_reinfections_90days_SIDIAP,
                  RR_first_infection_reinfections_90days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  correctSymptomNames()

 gg.rr.first_reinfection_90 <-  get_rr_first_reinfection_plot(data = RR_data)
gg.rr.first_reinfection_90

ggsave(file="Figure4.svg", plot=gg.rr.first_reinfection_90, width=10, height=8)
```

@@

### Incidence Rates
#### 1. Table: Incidence Rates of COVID-19 & Post-acute COVID-19 symptoms
```{r}

@@

plot_IR <- function(age_to_filter, data, brk_vec){ 
  plot <- data %>%
  filter(age_group %in% age_to_filter) %>%
    mutate(cohort = ifelse(cohort == "COVID-19 infections", "COVID-19 infections",
                           ifelse(cohort == "Long COVID symptoms after COVID-19 among the general population", "Post-acute COVID-19\nsymptoms among\nthe general population", "Post-acute COVID-19\nsymptoms among\nthose infected"))) %>%
  ggplot() +
#  geom_smooth(aes(x =incidence_start_date, y =  ir_100000_pys), method="auto", se=TRUE, fullrange=FALSE, level=0.95)+
geom_line(aes(x =incidence_start_date,
              y =  ir_100000_pys,
              color = age_group), size = 1.2)+
#    geom_ribbon(aes(x =incidence_start_date, 
    # y =  ir_100000_pys, color = age_group,
    # ymin = ir_100000_pys_95CI_lower, ymax = ir_100000_pys_95CI_upper
    # ), alpha = 0.1) +
#scale_y_continuous(limits = c(0,NA)) +
scale_x_date(breaks = brk_vec,
 #date_breaks = "2 month",
               date_labels = ("%b %y"),
               expand = c(0,0)) +
scale_colour_manual(values = c("#90e0ef","#00b4d8","#0077b6","#023E8A", "#01022F"
  #"#000000","#0075c4","#5b8c5a","#d8973c","#dc392e","#d8973c","#dc392e"
  # "#bee9e8","#62b6cb",","#cae9ff","#5fa8d3"
                                     )) +
    theme_bw()+
    theme(
      #panel.spacing.x = unit(4, "mm"),
      strip.text.y.left =  element_text(angle = 0),
      legend.position="top",
      plot.margin = unit(c(0, 0, 0, 0), "cm"))+
  ylab("IR per 100,000 p-y") +
 facet_grid(cohort ~sex  , 
            scales = "free") +    
    xlab("Date of COVID-19 infection") +
        guides(color = guide_legend(title = "Age group"))
   plot         
  }

@@

brk_vec <- seq.Date(as.Date("2020-10-01"),
                    as.Date("2021-12-01"),by="2 month")

@@

### Relative Risks - infections vs first negative test 
#### 1. Table: RR of Post-acute COVID-19 symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 90 days
```{r}
get_RR <- function(sidiap_data, cprd_data) {
  RR_infections <- rbind(sidiap_data, cprd_data
                      ) %>%
    filter(wave == "overall") %>%
    select(-"wave") %>%
  mutate(name_1 = "COVID-19 infection") %>%
  mutate(relative_risk = paste0(sprintf(relative_risk, fmt = '%.2f'),
                                " [",
                                sprintf(low_ci, fmt = '%.2f'),
                                ";",
                                sprintf(up_ci, fmt = '%.2f'),
                                "]"
         )) %>%
    select (-prop_1, -prop_2, -low_ci, -up_ci, 
                        -name_1, -name_2, -total_pop_1, -total_pop_2,) %>%
    relocate(database, .before = symptom)  %>%
    correctSymptomNames() %>%
    arrange(desc(database), desc(relative_risk))

@@

#### 2. Table: RR of Post-acute COVID-19 symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 28 days
```{r}
table_RR_28 <- get_RR(sidiap_data = RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  cprd_data = RR_new_covid_tested_negative_earliest_28days_CPRD)
table_RR_28
```

#### 3. Figure: RR of Post-acute COVID-19 symptoms: COVID-19 infection vs first SARS-CoV-2 negative test - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}
## Relative risks figure
 RR_data <- rbind(RR_new_covid_tested_negative_earliest_28days_SIDIAP,
                  RR_new_covid_tested_negative_earliest_28days_CPRD) %>%
  filter(wave == "overall") %>%
  select(-"wave") %>%
  correctSymptomNames()

@@

#### 2. Table: RR of Post-acute COVID-19 symptoms, first infection vs re-infection - 28 days
```{r}
table_RR_re_28 <- get_RR_numbers_re_infec(sidiap_data = RR_first_infection_reinfections_28days_SIDIAP,
                                       cprd_data = RR_first_infection_reinfections_28days_CPRD)
 table_RR_re_28
```

#### 3. Figure: RR of Post-acute COVID-19 symptoms, first infection vs re-infection - 28 days
```{r, fig.dim = c(8, 8), echo = FALSE}

RR_data <- rbind(RR_first_infection_reinfections_28days_SIDIAP,   RR_first_infection_reinfections_28days_CPRD) %>%
  filter(wave == "overall") %>% select(-"wave") %>%
  correctSymptomNames()
